---
title: "STA Final Project"
output: html_notebook
---
```{r}
library(tidyverse)
library(janitor)
dom_tour <- read_csv("data/regional_domestic_tourism.csv")
scot_accom <- read_csv("data/scottish_accomodation_occupancy.csv")
scot_activ <- read_csv("data/tourism_day_visits_activities.csv")
scot_demo <- read_csv("data/tourism_day_visits_demographics.csv")
scot_loca <- read_csv("data/tourism_day_visits_location.csv")
scot_tran <- read_csv("data/tourism_day_visits_transport.csv")

```


```{r}
dom_tour <- dom_tour %>%
  clean_names()

scot_accom <- scot_accom %>%
  clean_names()

scot_activ <- scot_activ %>%
  clean_names()

scot_demo <- scot_demo %>%
  clean_names()

scot_loca <- scot_loca %>%
  clean_names()

scot_tran <- scot_tran %>%
  clean_names()
```

# Cleaning the activity dataset

```{r}
scot_activ <- scot_activ %>%
  pivot_wider(names_from = units, values_from = value)

scot_activ <- scot_activ %>%
  select(-c(feature_code, measurement, breakdown_of_domestic_tourism))
  
```

# Revenue change over the years

```{r}
scot_activ %>%
  filter(tourism_activity == "All") %>%
  group_by(date_code) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
  geom_col(aes(x = date_code, y = revenue), size = 1, color = "darkblue", fill = "white") +
  scale_y_continuous(sec.axis = sec_axis(~./100))
```

# Visits Trend

```{r}
ggplot(scot_activ, aes(x = date_code)) +
  geom_col(aes(y = million_visits), size = 1, fill = "dark blue") 
  
```






# What kind of tourism activity generates the highest income for Scotland?

```{r}
scot_activ %>%
  filter(date_code == "2019" & tourism_activity != "All") %>%
  group_by(tourism_activity) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_col(aes(x = tourism_activity, y = revenue)) +
  scale_x_discrete(labels = c("All", "Spa Day", "Day Trip", "Entertainment", "Exploring", "Leisure Activities", "Night Out", "Outdoor Leisure", "Shopping", "Personal Events", "Public Events", "Sport Participation", "Visiting friends or family", "Visitor Attraction", "Live Sport", "Out for a meal"))
```
# Which kind of activity generates the most visits to sites?

```{r}
scot_activ %>%
  filter(date_code == "2019" & tourism_activity != "All") %>%
  group_by(tourism_activity) %>%
  summarise(visits = million_visits) %>%
  ggplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_col(aes(x = tourism_activity, y = visits)) +
  scale_x_discrete(labels = c("All", "Spa Day", "Day Trip", "Entertainment", "Exploring", "Leisure Activities", "Night Out", "Outdoor Leisure", "Shopping", "Personal Events", "Public Events", "Sport Participation", "Visiting friends or family", "Visitor Attraction", "Live Sport", "Out for a meal"))
```

# Is there a particular key demographic of visitors we should be focusing on? Who spends the most? Who visits the most?

```{r}
scot_demo <- scot_demo %>%
  select(-c(feature_code, measurement, breakdown_of_domestic_tourism))

scot_demo <- scot_demo %>%
  pivot_wider(names_from = units, values_from = value)

```

# Age category spends

```{r}
scot_demo %>%
  filter(date_code == "2019" & age != "All") %>%
  group_by(age) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
   geom_col(aes(x = age, y = revenue))
```
# Social status spends


```{r}
scot_demo %>%
  filter(date_code == "2019" & social_grade != "All") %>%
  group_by(social_grade) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
   geom_col(aes(x = social_grade, y = revenue))
```

# Employment status spends

```{r}
scot_demo %>%
  filter(date_code == "2019" & employment_status != "All") %>%
  group_by(employment_status) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
   geom_col(aes(x = employment_status, y = revenue))
```

# Child status

```{r}
scot_demo %>%
  filter(date_code == "2019" & children != "All") %>%
  group_by(children) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
   geom_col(aes(x = children, y = revenue))
```

# Access to car

```{r}
scot_demo %>%
  filter(date_code == "2019" & access_to_car != "All") %>%
  group_by(access_to_car) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
   geom_col(aes(x = access_to_car, y = revenue))
```

# 35-44 year olds seem to have the highest spend

```{r}
scot_demo %>%
  mutate(spend_per_visit = million_pounds_gbp/million_visits) %>%
  arrange(desc(spend_per_visit))
```

# Families with children seem to have a much higher spend than non 

```{r}
scot_demo %>%
  mutate(spend_per_visit = million_pounds_gbp/million_visits) %>%
  filter(children != "All") %>%
  arrange(desc(spend_per_visit))
```

# Find out which group has the most visits

16-24 years and then 25-34 years have most visits

```{r}
scot_demo %>%
  filter(age != "All") %>%
  arrange(desc(million_visits))
```

# What type of locations receive the most visits? What kinds of locations do people spend the most in?

```{r}
scot_loca <- scot_loca %>%
  select(-c(feature_code, measurement, breakdown_of_domestic_tourism))

scot_loca <- scot_loca %>%
  pivot_wider(names_from = units, values_from = value)
```

# revenue by location

```{r}
scot_loca %>%
  filter(date_code == "2019" & type_of_location_visited != "All areas") %>%
  group_by(type_of_location_visited) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
   geom_col(aes(x = type_of_location_visited, y = revenue))
```

# visits by location

```{r}
scot_loca %>%
  filter(date_code == "2019" & type_of_location_visited != "All areas") %>%
  group_by(type_of_location_visited) %>%
  summarise(visits = million_visits) %>%
  ggplot() +
   geom_col(aes(x = type_of_location_visited, y = visits))
```
# spend per visit per location

```{r}
scot_loca %>%
  filter(type_of_location_visited != "All areas") %>%
  mutate(spend_per_visit = million_pounds_gbp/million_visits) %>%
  arrange(desc(spend_per_visit))
```

# Is there a particular method of travel our visitors arrive by? Some of our locations are remote and not easily accessible by public transport, so this would be good information to have.

```{r}
scot_tran <- scot_tran %>%
  select(-c(feature_code, measurement, breakdown_of_domestic_tourism))

scot_tran <- scot_tran %>%
  pivot_wider(names_from = units, values_from = value)
```

# spend by method of travel

```{r}
scot_tran %>%
  filter(date_code == "2019" & method_of_travel != "All") %>%
  group_by(method_of_travel) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   geom_col(aes(x = method_of_travel, y = revenue)) +
  scale_x_discrete(labels = c("Bike", "Boat/Ship/Ferry", "Organised Tour", "Regular Bus", "Car Hired", "Car Owned", "Caravan/Campervan", "Lorry/Truck/Van", "Other", "Plane", "Taxi", "Train", "Tram", "Tube", "Walked"))
```
# Spend per visit by transport method
```{r}
scot_tran %>%
  mutate(spend_per_visit = million_pounds_gbp/million_visits) %>%
  arrange(desc(spend_per_visit))
```
# visits by method of travel

```{r}
scot_tran %>%
  filter(date_code == "2019" & method_of_travel != "All") %>%
  group_by(method_of_travel) %>%
  summarise(visits = million_visits) %>%
  ggplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   geom_col(aes(x = method_of_travel, y = visits)) +
  scale_x_discrete(labels = c("Bike", "Boat/Ship/Ferry", "Organised Tour", "Regular Bus", "Car Hired", "Car Owned", "Caravan/Campervan", "Lorry/Truck/Van", "Other", "Plane", "Taxi", "Train", "Tram", "Tube", "Walked"))
```

# When looking at regional tourism, what insights can we gain? Where are our visitors from? How do they differ in the money they spend, nights they spend in Scotland, or number of visits they do?


```{r}
dom_tour %>%
  filter(region_of_residence != "All of GB") %>%
  ggplot() +
   geom_col(aes(x = date_code, y = million_pounds_gbp, fill = region_of_residence)) 

```


```{r}
dom_tour %>%
  filter(region_of_residence != "All of GB") %>%
  ggplot() +
   geom_col(aes(x = date_code, y = thousand_nights, fill = region_of_residence))
  
```

```{r}
dom_tour %>%
  filter(region_of_residence != "All of GB") %>%
  ggplot() +
   geom_col(aes(x = date_code, y = thousand_visits, fill = region_of_residence))
```
# Highest spending region
```{r}
dom_tour %>%
  filter(region_of_residence != "All of GB") %>%
  arrange(desc(million_pounds_gbp))
```

# Highest number of nights

```{r}
dom_tour %>%
  filter(region_of_residence != "All of GB") %>%
  arrange(desc(thousand_nights))
```

# Spend per region

```{r}
dom_tour %>%
  filter(region_of_residence == "England" & date_code == "2015-2017") %>%
  sum(million_pounds_gbp)
```

# What type of accomodation is most popular? Is there a difference by location?

```{r}
scot_accom <- scot_accom %>%
  select(-c(feature_code, measurement, units))
```


```{r}
scot_accom %>%
  arrange(desc(value))
```

```{r}
scot_accom %>%
  filter(location != "All") %>%
  arrange(desc(value))
```

# Is there any way to predict spending and visits for Visit Scotland?


```{r}
library(tidyverse)
library(fable)
library(tsibble)
library(tsibbledata)
library(ggfortify)
library(ggplot2)
library(urca)
library(fpp2)
```

```{r}
forecast_tib <- as_tibble(forecast)
class(forecast_tib)

forecast_tib <- scot_activ %>%
  select(date_code, million_pounds_gbp)
```

```{r}
autoplot(as.ts(forecast_tib))
```

```{r}
glimpse(scot_activ)
```

# Splitting datset to date and revenue

```{r}
ts_data <- scot_activ %>%
  filter(tourism_activity == "All") %>%
  select(date_code, million_pounds_gbp)

ts_data_nas <- na.omit(ts_data)

y <- ts(ts_data_nas[,2],start=c(2013),frequency = 1)
```

# Preliminary Analysis - Time Plot

```{r}
autoplot(y) 


```

```{r}
fit <- snaive(y)
print(summary(fit))
checkresiduals(fit)
```

```{r}
forecast_1 <- fit %>%
  fabletools::forecast(h = 1)
forecast_1
```

```{r}
forecast_1 %>%
  autoplot(y) 
  library(RColorBrewer)
```


```{r}
scot_activ %>%
   filter(tourism_activity == "All") %>%
  ggplot() +
geom_col(aes(x= date_code, y = million_pounds_gbp)) +
  labs(
    x = "Year",
    y = "Million Pounds Revenue",
    title = "Scottish Tourism Activity Revenue 2013 - 2019"
  )
```

```{r}
scot_activ %>%
  filter(date_code == "2013" & tourism_activity == "All")
```

