---
title: "Scottish Tourism Report"
author: "Graham Angus"
date: "10/03/2021"
output: html_document
---

#Analysis of The Visit Scotland Dataset
## By Graham Angus

**Intro to the data**

My final project is based on the Visit Scotland dataset which comprises 6 datasets 
to answer the below questions and any more detailed ones. 

• What kind of tourism activity generates the highest income for Scotland?
• Which kind of activity generates the most visits to sites?
• Is there a particular key demographic of visitors we should be focusing on? 
Who spends the most? Who visits the most?
• What type of locations receive the most visits? 
What kinds of locations do people spend the most in?
• Is there a particular method of travel our visitors arrive by? 
Some of our locations are remote and not easily accessible by public transport, 
so this would be good information to have.
• When looking at regional tourism, what insights can we gain? 
Where are our visitors from? How do they differ in the money they spend, 
nights they spend in Scotland, or number of visits they do?
• What type of accomodation is most popular? Is there a difference by location?
• How have general tourism rates changed over time?
• Is there any way to predict spending and visits for Visit Scotland?

The datasets are made up from several different methods, the regional tourism one
being Interviewing is carried out face-to-face every week of the year. 
Annually, some 100,000 respondents are contacted, and any who have returned from 
an overnight trip within the past four weeks are asked to describe the details of 
that trip.

The scottish accomodation data is taken from SAOS is a voluntary monthly survey 
and participation of accommodation providers varies across type and time.
This means that the survey only allows for occupancy rate data which can make 
full analysis difficult. 

The other 4 datasets are derived from The Great Britain Day Visits Survey started 
in 2011 and was conducted using an online methodology. Over 35,000 interviews 
are conducted with adults aged 16 and over resident in England, Wales and 
Scotland during 52 weekly survey waves. This is an opt in online questionnaire 
which can come with issues.

**The Domestic Tourism datset**

```{r}
country_dom_tour %>%
  filter(breakdown_of_domestic_tourism == "Visits") %>%
  ggplot()+
   theme(legend.position='none') +
  ggtitle("Scottish Tourist Region of Residency 2019") +
   ylab("Visits per Thousand") +
  xlab("Year") +
  geom_col(aes(x = date_code, y = value, fill = region_of_residence))
```

I was able to find out that domestic tourism rates have stayed largely flat over
the period and the number from England and Scotland is relatively similar 
especially when you take the differing populations into context. 



**Scottish Accomodation dataset**

This is probably the trickiest dataset as it is firstly based on % occupancy which 
isn't a great way to gauge success as it doesn't tell you wether supply has 
increased or decreased and likewise with demand. There are wild variances in the
data which go against what would be expected and the sorting of the survey data 
is to be one of my recommendations.


**Scottish Activity Datset**

This dataset gives us an insight into what activites tourists are coming to do in
Scotland and also what they have done once here. The following graphs show a 300% 
increase in revenue from personal events like weddings, a doubling of live sport
revenue from tourists and a 25% increase in meals out revenue. Generally there is 
an increase in all of leisure/sport type categories and no change in the Visitor 
Attractions revenue. 

```{r}
scot_activ %>%
  filter(date_code == "2013" & tourism_activity != "All") %>%
  group_by(tourism_activity) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_col(aes(x = tourism_activity, y = revenue, fill = tourism_activity)) +
   ggtitle("Scottish Tourist Activity Revenue 2013") +
   ylab("Revenue per Million") +
  xlab("Tourist Activities") +
  theme(legend.position='none') +
  scale_x_discrete(labels = c("Spa Day", "Day Trip", "Entertainment", "Exploring", "Leisure Activities", "Night Out", "Outdoor Leisure", "Shopping", "Personal Events", "Public Events", "Sport Participation", "Visiting friends or family", "Visitor Attraction", "Live Sport", "Out for a meal"))
```

```{r}
scot_activ %>%
  filter(date_code == "2019" & tourism_activity != "All") %>%
  group_by(tourism_activity) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_col(aes(x = tourism_activity, y = revenue, fill = tourism_activity)) +
   ggtitle("Scottish Tourist Activity Revenue 2019") +
   ylab("Revenue per Million") +
  xlab("Tourist Activities") +
  theme(legend.position='none') +
  scale_x_discrete(labels = c("Spa Day", "Day Trip", "Entertainment", "Exploring", "Leisure Activities", "Night Out", "Outdoor Leisure", "Shopping", "Personal Events", "Public Events", "Sport Participation", "Visiting friends or family", "Visitor Attraction", "Live Sport", "Out for a meal"))
```




**Scottish Demographic datset**

This dataset gives us insight into age, marital status, social class, gender, 
employment status, access to car and children. It is another difficult dataset 
to wrangle or even model as each column has an all value which encompasses 
every cateogry in the variable so stops further analysis but I was able to split
and find variations in age brackets and average spend and the change over the years.

```{r}
scot_demo %>%
  filter(age != "All" & date_code == "2019") %>%
   group_by(age) %>%
   summarise(avg_spend = million_pounds_gbp/million_visits)
```


```{r}
scot_demo %>%
  filter(age != "All" & date_code == "2013") %>%
   group_by(age) %>%
   summarise(avg_spend = million_pounds_gbp/million_visits)
```


**Scottish Location dataset**

This dataset looks at the location tourists visit whilst in Scotland for day 
activities, I feel this may skew the data as I feel people may be more likely to
visit bigger towns and cities for activities even if they are staying outside of 
them. The graph below looks at revenue and visits over time and you can see Cities 
has increased over time where villages and more rural locations have stayed 
broadly static. 

```{r}
scot_loca %>%
  ggplot() + 
  ggtitle("Scottish Tourist Locations 2019") +
   ylab("Visits per Million") +
  xlab("Revenue per Million") +
  geom_point(aes(x = million_pounds_gbp, y = million_visits, colour = date_code))
```



**Scottish Transport dataset**

This dataset shows the type of transport used by tourists whilst here, there are 
numerous findings which don't make intuative sense and need further explanation
Some of the findings for example show 
Bus journeys have stayed at same level but income has dropped by ⅓
Internal Flights 0.2m to 1.3m
Bike 1.5m to 1.2m visits but income £3.4m to £15.5m
Ferry 0.8m to 0.5m visits but income £12.2m to £27.7m
Campervan usage static but income £5.1m to £24.4m



```{r}
scot_tran %>%
  filter(date_code == "2013" & method_of_travel != "All" & method_of_travel != "Car - own/friends/family" & method_of_travel != "Motor cycle"
          & method_of_travel != "Bus or Coach - Minibus") %>%
  group_by(method_of_travel) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   geom_col(aes(x = method_of_travel, y = revenue, fill = method_of_travel)) +
    ggtitle("Scottish Transport Methods 2013") +
  ylab("Revenue £m") +
  xlab("Method of Travel") +
    scale_y_continuous(breaks = c(200, 400, 600, 800)) +
  theme(legend.position='none') +
  scale_x_discrete(labels = c("Bike", "Boat/Ship/Ferry", "Organised Tour", "Regular Bus", "Car Hired", "Caravan/Campervan", "Lorry/Truck/Van", "Other", "Plane", "Taxi", "Train", "Tram", "Tube", "Walked")) 
```




```{r}
scot_tran %>%
  filter(date_code == "2019" & method_of_travel != "All" & method_of_travel != "Car - own/friends/family") %>%
  group_by(method_of_travel) %>%
  summarise(revenue = million_pounds_gbp) %>%
  ggplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   geom_col(aes(x = method_of_travel, y = revenue, fill = method_of_travel)) +
    ggtitle("Scottish Transport Methods 2019") +
   ylab("Revenue £m") +
  xlab("Method of Travel") +
  scale_y_continuous(breaks = c(200, 400, 600, 800)) +
  theme(legend.position='none') +
  scale_x_discrete(labels = c("Bike", "Boat/Ship/Ferry", "Organised Tour", "Regular Bus", "Car Hired", "Caravan/Campervan", "Lorry/Truck/Van", "Other", "Plane", "Taxi", "Train", "Tram", "Tube", "Walked"))
```



**Conclusions**

Data is never going to be perfect but these particular sets don't allow for 
further analysis and to asnwer questions about wether there is higher demand and 
supply for accomodation, for further exploration into demographics which could 
help with marketing targeting and also service provision. It seems like tourism 
revenue has increased over the period and some metrics like campervans anecdotally 
do seem to more prominent in Scotland which can often come with problems, in this
specific case a plan on how to house and accomodate campervans within our 
infrastructure is much needed and would possibly help those affected by them.